An Open Platform for Managing Short Programming Exercises

### 2013-06-25 paper/paper.tex
minor edit

Recent years have seen an explosion of online services for learning
to program:  Udacity\cite{udacity}, Codecademy\cite{codeacademy},
ProgZoo\cite{progzoo}, 
CodingBat\cite{codingbat}, Coursera\cite{coursera}, CodeLab\cite{turingscraft},
Problets\cite{Kumar:2005:GPA:1163405.1163408}, CS Circles\cite{Pritchard:2013:CCI:2445196.2445370} and
many others.  These systems are powerful, and show great promise to
improve introductory computer science.  Given the existence of all of these
systems, one may ask why we thought
%it a good career move
%to build yet another system.
the world needed another such system.

@@

Figure~\ref{screenshot} shows a screenshot of the user interface
students use to work on a CloudCoder exercise.
CloudCoder uses Ace\cite{ace}, an open source web-based
programmer's editor, which supports
syntax highlighting and automatic indentation for most programming languages.
The submit button runs the code against a series of
test cases. Once the test results are received, they are presented
directly in the browser, giving the student immediate feedback on the
correctness of the submission.
CloudCoder captures the code's standard output and
standard error streams to help students with debugging.

@@

CloudCoder is available under the GNU Affero General Public License\cite{AGPL}.
Binary distributions are available through our website (\url{http://cloudcoder.org}),
and the code is available on GitHub:
\begin{quote}
\url{https://github.com/cloudcoderdotorg/CloudCoder}
\end{quote}

@@

Whole-program exercises read input data from
the standard input stream, perform a computation, and write output to the standard output stream.
Expected output is checked by matching a regular expression against the
program's standard output stream.
Whole-program exercises are useful for two main reasons.  First,
they can be assigned early in the course, before functions/methods
have been introduced.  Second, they give the exercise author a
way of checking expected output that is more powerful than matching a return
value against a literal.  The main disadvantage is
the difficulty of devising a
regular expression that is precise enough to distinguish correct output
from incorrect output, but forgiving enough to allow minor deviations
in white space, capitalization, and formatting of output.

@@

{\bf How can programming exercises be integrated most effectively into a course?}
Programming exercises can be used for many purposes: self-test questions
to accompany readings and video lectures, in-class assessment (possibly
in the context of peer instruction), extra practice for students who
are struggling.  We would like to have specific, evidence-based recommendations
that we can give to instructors who are considering using programming
exercises in their courses to guide them in using exercises in an effective manner.

@@

{\bf What is the right balance between short programming exercises and
  longer ``nifty'' assignments?}
There has been a lot of work in the CS Education community developing
interesting and innovative introductory level programming assignments,
such as the wonderful examples presented each year at the ``Nifty Assignments'' panel at
SIGCSE.  Most introductory courses require at least a few
longer nifty-style programming assignments, in addition to shorter
programming homework assignments.  What is the proper balance between longer
nifty-style assignments and shorter, more focused programming exercises?  Is 
solving shorter programming exercises a pre-condition, or threshold skill, for
completing assignments?

{\bf Is the feedback provided by synchronous grading more effective
  than feedback from asynchronous grading?}
Faculty or Teaching Assistants who manually grade student programming
assignments can give detailed and customized feedback to
students, at the cost of some latency between the time the student
submits the assignment and the time they receive their grade and the
feedback.  Anecdotally, many faculty complain that students do not
seem to pay much attention to this feedback, possibly because those
points (marks) have already been lost and cannot be recovered.
While automated systems cannot give feedback that is as detailed as manual
grading, they can give feedback very quickly.

@@

\section{Acknowledgments}

@@

This work was supported by a SIGCSE Special Projects grant (May 2012),
a York College Summer Research Fellowship (Summer 2013), an award from
the Paul K Richter and Evalyn E. Cook Richter Memorial Trusts at Knox
College, and an Amazon Web Services Teaching Grant.

>>>>

Recent years have seen an explosion of online services for learning
to program:  Udacity\cite{udacity}, Codecademy\cite{codeacademy},
ProgZoo\cite{progzoo}, 
Codingbat\cite{codingbat}, Coursera\cite{coursera}, CodeLab\cite{turingscraft},
Problets\cite{Kumar:2005:GPA:1163405.1163408}, CS Circles\cite{Pritchard:2013:CCI:2445196.2445370} and
many others.  These systems are powerful, and show great promise to
improve introductory computer science.  Given the existence of all of these
systems, one may ask why we thought
%it a good career move
%to build yet another system.
the world needed another such system.

@@

Figure~\ref{screenshot} shows a screenshot of a CloudCoder exercise.
CloudCoder uses Ace\cite{ace}, an open source web-based
programmer's editor, which supports
syntax highlighting and automatic indentation for most programming languages.
The submit button runs the code against a series of
test cases. Once the test results are received, they are presented
directly in the browser, giving the student immediate feedback on the
correctness of the submission.
CloudCoder captures the code's standard output and
standard error streams to help students with debugging.

@@

CloudCoder is available under the GNU Affero General Public License\cite{AGPL}.
Binary distributions are available through our website (\url{http://cloudcoder.org}),
and the code is available on Github:
\begin{quote}
\url{https://github.com/cloudcoderdotorg/CloudCoder}
\end{quote}

@@

Whole-program exercises read input data from
the standard input stream, perform a computation, and write output to the standard output stream.
Expected output is checked by matching a regular expression against the
program's standard output stream.
Whole-program exercises are useful for two main reasons.  First,
they can be assigned early in the course, before functions/methods
have been introduced.  Second, they give the exercise author a
way of checking expected output that is more powerful than matching a return
value against a literal.  The main disadvantage is
the difficulty of devising a
regular expression that is precise enough to distinguish correct output
from incorrect output, but forgiving enough to allow minor deviations
in whitespace, capitalization, and formatting of output.

@@

{\bf How can programming exercises be integrated most effectively into a course?}
Programming exercises can be used for many purposes: self-test questions
to accompany readings and video lectures, in-class assessment (possibly
in the context of peer instruction), and extra practice for students who
are struggling.  We would like to have specific, evidence-based recommendations
that we can give to instructors who are considering using programming
exercises in their courses to guide them in using exercises in an effective manner.

@@

{\bf What is the right balance between short programming exercises and
  longer ``nifty'' assignments?}
Discuss.

{\bf What can we do to encourage students to use their time effectively outside of class?}

@@

\section{Acknowledgements}

@@

This work was supported by a SIGCSE Special Projects grant (May 2012),
a York College Summer Research Fellowship (Summer 2013), [TODO: Knox College
grant], and [TODO: Amazon grant].
